---
output: github_document
---

## Querying a GraphQL client for linked data using R

<br />

The following script relies on the [`ghql`](https://github.com/ropensci/ghql) R package (Chamberlain 2017) to query multidimensional QB datasets using [GraphQL](http://graphql.org/). The example uses the graphql-qb service at [graphql-qb.publishmydata.com](graphql-qb.publishmydata.com) which stores data from [statistics.gov.scot](statistics.gov.scot).

---

##### Install the `ghql` package from GitHub (and `devtools` if not already installed)
```{r, eval=FALSE}
devtools::install_github("hadley/devtools")
devtools::install_github("ropensci/ghql")
```

##### Load the necessary R packages
```{r, message=FALSE, warnings=FALSE}
library(ghql) # for querying 
library(jsonlite) # for parsing the json response
library(httr) # for working with URLs
library(tidyverse) # for tidying data
```

##### Initialize the GraphQL client by pointing it to the appropriate endpoint e.g. http://graphql-qb.publishmydata.com/graphql
```{r}
client <- GraphqlClient$new(url = "http://graphql-qb.publishmydata.com/graphql")
```

No OAuth token is required for this endpoint but the `headers` argument can be used for this purpose.

##### Make a Query class object
```{r}
qry <- Query$new()
```

##### Add your GraphQL query, e.g. Filter datasets about gender and return the title and description
```{r}
qry$query('query', '
  {
    datasets(dimensions: {and: ["http://statistics.gov.scot/def/dimension/gender"]}) {
      title
      description
    }
  }
')
```

##### Return the responses
```{r}
responses <- client$exec(qry$queries$query)
```

##### Convert to a dataframe and return column names
```{r}
df <- as.data.frame(responses)
glimpse(df)
```

##### Change the column names
```{r}
df <- rename(df, Dataset = data.datasets.title,
             Description = data.datasets.description)
```

The table below shows the first 6 responses.

```{r, echo=FALSE}
library(knitr) ; library(kableExtra)
kable(head(df), format = "html") %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

<br />

### References

- Chamberlain, Scott (2017). ghql: General Purpose GraphQL Client. R package version 0.0.3.9110. [https://github.com/ropensci/ghql](https://github.com/ropensci/ghql)      
- Ooms, Jeroen (2014). The jsonlite Package: A Practical and Consistent Mapping Between JSON Data and R Objects. arXiv:1403.2805 [stat.CO] URL [https://arxiv.org/abs/1403.2805](https://arxiv.org/abs/1403.2805).  
- Wickham, Hadley (2017). httr: Tools for Working with URLs and HTTP. R package version 1.3.1. [https://CRAN.R-project.org/package=httr](https://CRAN.R-project.org/package=httr)  
- Wickham, Hadley (2017). tidyverse: Easily Install and Load 'Tidyverse' Packages. R package version 1.1.1. [https://CRAN.R-project.org/package=tidyverse](https://CRAN.R-project.org/package=tidyverse)

